Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2017-08-11 sections0.08_for_editor/results0.08.tex
re-submitted version

    In this section we show the results of the neural networks and K-means clustering trained using the \citetalias{Kinney96} template spectra.
    Training with \citetalias{Kinney96} templates results in networks/clusters that have regions corresponding to galaxies of known morphological type. 
    The trained networks/clusters can then be used to categorize other galaxies.

@@

            Figs.~\ref{fig: 1by2T} --~\ref{fig: 1by22T} show the results of the training networks.  %(the remaining networks are shown in Appendix~\ref{app: high_Z_1d_soms}).
            The upper part of the figures shows the neurons and the relative distances between their weights. 
            As mentioned in Section~\ref{sec: method_somz}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.45\textwidth}
                    \centering
                  \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.45\textwidth}
                    \centering \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption[Results of training network in $1\times2$~grid]{Results of training network in $1\times2$~grid. The upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citet{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

@@

    We performed K-means clustering on the \citetalias{Kinney96} template spectra using 4 clusters.
    \textbf{
    Fig.~\ref{fig: som_k_means_4} shows the clustering of the template spectra into 4 groups using both K-means and SOMs.
    Both methods separated the most starburst-like and quiescent galaxies in a similar manner.
    However, they show discrepancies for spectral types between these extreme cases.
    Table~\ref{tab:kmeans_som4} shows the group assignments of the \citetalias{Kinney96} templates in the two methods.
    Templates B, E, S0 and Sa are grouped in the same cluster/neuron in both K-means and $1\times4$-sized SOM. 
    K-means grouped the Sb template with the above templates, but the SOM network classified it as a separate group.
    The K-means method clustered the Sc template as a separate group, while the SOM network grouped it with SB5 and SB6. 
    Another noticeable difference is that K-means clustered the SB1 and SB2 templates in one group and SB3 to SB6 in another while
    the SOM network grouped the SB1 to SB4 templates in one neuron and SB5 and SB6 in another.
    As Fig.~\ref{fig: som_k_means_4} shows, K-means clustering of the templates produced groups more similar to each other, while SOM emphasizes the templates' differences.
    }

@@

%Sahar: here is where we discuss which of the templates is in each group for the 2 methods. Possibly also introduce the performance metric here.

@@

         \textbf{After training with the templates, we used both K-means and self-organizing maps to classify the 142 galaxies in the \citetalias{Hossein12} sample.
         K-means classification was performed by assigned each galaxy to the K-means cluster with the closest centroid and the results are discussed in Section~\ref{sec: Kmeansvssomvsann}.
        This section examines the results of one-dimensional SOM-based classification in detail.
         }

@@

            Since in this network galaxies were forced to be divided into a maximum of two groups, the strongest feature in a galaxy's spectrum predominantly decided to which group the galaxy belonged.
            Galaxies with a weak 4000\AA~break, strong emission lines and ultraviolet upturn are categorized as starbursts while galaxies with a strong 4000\AA~break are categorized as quiescent.
            Increasing the size of the SOM helps solve the problem of galaxies which have features that are common in both groups.

            Fig.~\ref{fig: 1by3V} presents the result of classifying the spectra of the galaxies using the $1\times3$~network (from Fig.~\ref{fig: 1by2T}): 66 of the galaxies belong to the quiescent group, and 47 belong to the starburst group. 
            However, 29 of the galaxies are similar to starbursts in some, but not all, of their features. 
            Galaxies in this group have strong emission lines and are ultraviolet-bright, but they also have a strong 4000\AA~break, which makes them cluster closer to the quiescent galaxies (middle panel in the lower part of Fig.~\ref{fig: 1by2V}).

@@

    \subsubsection{Properties of the spectral groups}

       In order to check whether our clustering in Sec.~\ref{sec: 1Dv}~is meaningful, in this section we examine the relations between median properties of the galaxies in each group. 
          Fig.~\ref{fig: props} shows the median of stellar mass, age, specific star formation rate (sSFR; star formation rate per stellar mass), and $f_\mathrm{burst}$ of the galaxies in each neuron in the $1\times2$, $1\times4$, and $1\times8$ networks.
        In all plots, the horizontal axis is the number of the neurons divided by the size of the network.
        As shown in Fig.~\ref{fig: props}, in all three networks, stellar mass and age decrease while sSFR and $f_\mathrm{burst}$ increase as the type changes from quiescent galaxies to starburst.
       Separating galaxies based on spectral types also leads to a separation in properties derived (via {\em CIGALE}) from the SEDs, as expected since the spectral types are also based on {\em CIGALE} fitting.

        \begin{figure*}
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f1.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f2.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f3.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f4.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f5.png}
        \end{subfigure}
       \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f6.png}
        \end{subfigure}
        \caption{From top to bottom: relations between age (t$_{\rm {D4000}}$) and sSFR ($\phi$), sSFR and stellar mass, age (t$_{\rm {D4000}}$) and stellar mass, and A$_{\rm {FUV}}$ with sSFR, age (t$_{\rm {D4000}}$) and stellar mass. Square points show the median values of these properties of galaxies in the $1\times22$ SOM network and error bars are the standard deviation of the median values. Circles show the median value of properties inside each K-means cluster (using $K=4$) and asterisks show results from \citetalias{Hossein12}. Colours show the galaxy types in the SOM classification. The lightest grey (most pure red in the online version) colour shows E type galaxies and the black (most blue in the online version) indicates the SB1 galaxies. The shading of other colours between light grey and black (red and blue in the online version) measures the location of galaxies along the SOM from  quiescent to starburst.}
        \label{fig: props_vs_props}
        \end{figure*}

        In order to compare our classifications with previous work, we produced similar plots to the ones in both \citetalias{Noll09} and \citetalias{Hossein12} (Fig.~\ref{fig: props_vs_props}).
        In Fig.~\ref{fig: props_vs_props}, the point represent the median values of the properties of the galaxies in each group, with different symbol types representing classification method and error bars indicating the standard deviation of those properties.
       The colour scale represents neuron number in the SOM method, shading from quiescent to starburst. 
        Since in the $1\times22$-size SOM shown in Fig.~\ref{fig: 1by22V} we allow galaxies to occupy classes in between the original \citetalias{Kinney96} templates, there are more than 12 colours in Fig.~\ref{fig: props_vs_props}.
        The medium grey (purple in the online version) point in the middle of the black (blue in the online version) points in Fig.~\ref{fig: props_vs_props} corresponds to galaxies with spectra similar to type Sc: the shape of their spectra indicates that they are old galaxies with high sSFR.

        The upper panels in Fig.~\ref{fig: props_vs_props} show the same relations as \citetalias{Noll09} and \citetalias{Hossein12}, but with more precision than those two, in all three plots.
        % *** MORE PRECISION - NEED TO CLARIFY
        As mentioned in both \citetalias{Noll09} and \citetalias{Hossein12}, galaxies with smaller mass tend to be younger and more active.
        Fig.~\ref{fig: props} shows similar results: younger and more active galaxies have lower stellar mass and higher f$_{\rm {burst}}$.
        From older to younger galaxies, the colour of the points changes from light grey to black (red to blue in the online version), which shows a good correlation with their spectral type.

        \citetalias{Noll09} studied relations between A$_{\rm {FUV}}$ and other properties of galaxies and showed that the attenuation has no dependence on the specific star formation or age.
        In contrast to \citetalias{Noll09}, we find a correlation between A$_{\rm {FUV}}$ and these two parameters, shown in the lower panels of Fig.~\ref{fig: props_vs_props}.
        The general trend of the correlation between A$_{\rm {FUV}}$ and sSFR in Fig.~\ref{fig: props_vs_props} is similar to the trend found by \cite{Dale07}.
        They used IR luminosity to UV luminosity ratio (L$_{IR}$/L$_{FUV}$) as a measure of A$_{\rm {FUV}}$ and compared it with sSFR for all 75 galaxies in the SINGS survey.
        They found that for quiescent galaxies (E, S0 and S0/a), L$_{IR}$/L$_{FUV}$ (or A$_{\rm {FUV}}$) correlates with sSFR, and for spiral galaxies, there is an anticorrelation between L$_{IR}$/L$_{FUV}$ and sSFR.
        Since our sample has no E-type galaxies, we cannot confidently show the same correlation between A$_{\rm {FUV}}$ and sSFR. 
        However, it is clear in Fig.~\ref{fig: props_vs_props} that A$_{\rm {FUV}}$ increases with increasing sSFR in the S0 and Sa types.
        The correlation in the other types of galaxies is very similar to the one shown by \cite{Dale07}.
        Both \cite{Dale07} and \citetalias{Noll09} argued that these apparent trends can be a result of the dependence of star formation history on L$_{IR}$/L$_{FUV}$.
        Whether this dependence is real or not, our results here show that using self-organizing maps can separate spectra of galaxies in such a way that the characteristics of each of these groups are in agreement with the general picture of the galaxy evolution.

       \textbf{One of the key features of SOMs is the ability to not only classify objects but to provide an {\em ordered} classification.
       K-means clustering and supervised ANNs can sort galaxy spectra into classes but do not naturally order them. Ordering is important in the case where the physical causes of the differences between classes/clusters are unknown. Average properties of objects within clusters (possibly from data not used for classification) can be used to test hypotheses about the physical differences. 
       Figures~\ref{fig: props} and~\ref{fig: props_vs_props} show that the SOM ordering also corresponds to the ordering of the physical properties of the galaxy models used to make the classified spectra. 
       The median properties of the spectral groups change on going from one end of the SOM to the other.
       Therefore, by using the SOM method, we can study not only how galaxies group into spectral clusters, but the trends between the clusters as well.}

@@

        As described in Section~\ref{sec: method_somz}, one of the main advantages of the SOM is that when the weight of one neuron is adjusted after finding the best matching unit, the weight of the whole map will be changed.
        This quality of the SOMs provides a unique opportunity to analyse 2D networks using two approaches. 
        At first, as in the 1D SOMs, we assumed that all the galaxies must have spectra similar to any of the 12 templates in Fig.~\ref{fig: k96}~(i.e. no outliers).
        In this case, we can categorize spectra of other galaxies based on those 12 types.
        In the second approach, we trained networks with the assumption that there might be completely different types of galaxies not encountered before.
        This approach can be used to trace and recognize the outliers, which could include galaxy types that are not present in the \citetalias{Kinney96} templates (e.g. dwarf galaxies, luminous infrared galaxies, or active galactic nuclei).
        To execute the second approach, we used the \textsc{ selforgmap} library in \textsc{matlab}.
        Unlike the \textsc{newsom} library in which the ordering step neighbourhood distance is set by default (although in an SOM it should be a free parameter), in \textsc{selforgmap} we can change this value.
        Using the \textsc{selforgmap} code, based on the size of a map and the ordering step neighbourhood distance, we can arrange that the data spread all over the SOM or sit close together, allowing room for outliers.

@@

        In the upper left part of the map in Fig.~\ref{fig: 12by12}, although galaxies have more ways to be separated, they were separated into two main groups.
        There is a distinguishable strip of grey, dark grey and in some cases black colour in the map:
        this strip separates quiescent galaxies from starburst ones.
        The lower map shows that 5 of the neurons on the left side of the strip are full. 
        These five neurons are the same as the ones on the left hand side of Fig.~\ref{fig: 1by2T} (quiescent galaxies),
        the only difference being that in this map they have more space to be separated from each other.
        When we use this network to categorize the spectra of galaxies, any galaxy placed on the left hand side of the strip is a quiescent galaxy and any one placed on the right side of the strip is a starburst one.
        The decision about what sub-type of quiescent or starburst galaxy is based upon its position relative to each type in the SOM.
        \citetalias{Hossein12} showed that by masking emission lines (i.e., by ignoring details in spectra) two types of galaxies, Sb and SB6, can be misclassified as one another.
        As mentioned before, few-node, one-dimensional SOMs are a good method to obtain a big picture of the problem under study with less emphasis on fine detail. 
        For example, the SB6 category is placed near Sb galaxies in the 1D network in Fig.~\ref{fig: 1by22T}.
        However, in the upper left panel of Fig.~\ref{fig: 12by12}, we can see that the Sb and SB galaxies are completely separated. 
        This shows that more details of spectral shapes are considered by 2D networks, which is one of the main advantages of  using 2D maps over 1D ones.
        \textbf{In this case, the two dimensions of the self-organizing map correspond approximately to quiescent versus star-forming (left to right) and less to more extinction (bottom to top).}

@@

        We used both 2D networks to categorize the \citetalias{Hossein12} galaxies and the results are shown in the lower panels of Fig.~\ref{fig: 12by12}.
        In the lower right map in Fig.~\ref{fig: 12by12}, all galaxies are placed in the bottom part of the map and we can conclude that in the \citetalias{Hossein12} sample there are no outliers or spectra with very different from the \citetalias{Kinney96} templates.
        The lower left map of Fig.~\ref{fig: 12by12} shows the \citetalias{Hossein12} galaxies categorized based on the network in the upper left of Fig.~\ref{fig: 12by12}. 
        Comparing this categorization with the 1D one from Fig.~\ref{fig: 1by22V}, we can see that only 23 galaxies correspond exactly to \citetalias{Kinney96} types.
        Using 2D maps results in categorizing galaxies into more intermediate types than in 1D maps.
        In both lower maps in Fig.~\ref{fig: 12by12}, most galaxies are in the quiescent side of the SOM, which was predictable from the results in Section~\ref{sec: 1Dv}.
        This does not \textbf{necessarily} imply that in general there are more quiescent galaxies at higher redshift but could be a selection effect for galaxies that had more reliable redshift estimates.

    Although for ease of presentation this paper first discusses 1D networks and continues to 2D networks, we suggest that users of SOMs for galaxy spectral classification should use 2D networks first.
         These can be used as a first step with networks similar to the upper left map in Fig.~\ref{fig: 12by12} to identify outliers in the sample.
        After finding and removing outliers in the input data, the use of 1D and 2D maps is complementary.
        One-dimensional maps are helpful for exploring the general behaviour of the data, while 2D maps show a more detailed picture.
        % Sahar: can we say anything more quantitative here about the difference betwen 1D and 2D?

\subsection{Comparison of the self-organizing map, K-means, and supervised ANN methods}
    \label{sec: Kmeansvssomvsann}

    The one and two-dimensional self-organizing map classifications are compared briefly above. 
    The 2-D maps allow galaxies to be members of more intermediate classes, and this is a general advantage of SOMs compared to the artificial neural network method used by \citetalias{Hossein12}: in the SOM method galaxies do not need to exactly match template spectra. 
    \citetalias{Hossein12} found that the ANN method could only match 105 out of the 142 galaxies directly to the \citetalias{Kinney96} templates; classifying the remaining 37 galaxies required combining template spectra.

%From introduction:     
%The questions we want to answer are: 
%(1) What is the degree of agreement between $\chi^2$ fitting, one- and two-dimensional self-organizing maps, K-means clustering and artificial neural networks? 
%(2) Are certain spectral types more (or less) consistently classified by the different methods? 
%(3) To what extent do the average properties of galaxy classes represent a continuous distribution of physical properties?

        \textbf{In order to compare results from SOM, K-means, and unsupervised ANN, we used two different methods.
        We assigned the 142 galaxies in the sample to the best fit \citetalias{Kinney96} template using minimum chi-square.
        We scored the classifications by K-means, $1\times4$-sized SOM, $1\times12$-sized SOM $1\times22$-sized SOM and unsupervised ANN using the following system.
        %Sahar: I rewrote this to make it clearer, but I'm not sure it's correct!!!
        If a galaxy's minimum chi-square template was classified into the same cluster/neuron, this scored 1 point. 
        If the galaxy was classified between two nodes (in the case of $1\times12$-sized SOM and $1\times22$-sized SOM) this scored 0.5 point. 
        Otherwise, the classification score was zero points. 
        Therefore, if the total score of a galaxy is 5, it means that all techniques classified the galaxy into the same group.
        If it is 4.5, it means that either in the 12-node SOM  or the 22-node SOM, that galaxy belonged to the in-between area and so on.
        Comparing the scores, we found that 12 per cent of galaxies scored 5, 4 per cent had a score of 4.5 and 20 per cent had a score of 4.
        In total, 36 per cent of galaxies had a score greater than 4 and 56 per cent had a total score of 3 or more.
        The score$=5$ galaxies classified similarly by all methods were primarily (60 per cent) in the Sa template group. %; the rest belonged to the Sb, SB1, SB4 and SB6 template groups.
        The score $>=3$ galaxies were predominantly SB1, Sa, and Sb type (32, 28, 18 per cent, respectively).
        The different methods show the most agreement for galaxies similar to the SB1 and Sa templates, with less agreement on the high-extinction and quiescent types (SB5, SB6).
        }

        \textbf{To see which methods gave the most similar results to the minimum chi-square fitting method, we summed the scores over all galaxies for each of the methods and found totals of K-means 124, $1\times4$-sized SOM 103, $1\times12$-sized SOM 111.5, $1\times22$-sized SOM 79, and unsupervised ANN 52.
        This shows that the K-means method and $1\times12$-sized SOM have the most similarity with minimum chi-square fitting.
        However, none of the methods produced results that were extremely similar to chi-squared
        }

        \textbf{To check the degree of agreement between different methods, we measured the Fleiss kappa index. 
        The Fleiss Kappa index is a measure of agreement between three or more raters that are classifying a specific number of objects (i.e. 142 galaxies) to a number of classes (i.e. 12 templates).
        We computed two different values of Fleiss kappa, between  K-means, unsupervised ANN, and either 
         $1\times12$-sized or $1\times22$-sized SOMs.
        The resulting values were $\sim 27$ per cent and $\sim 21$ per cent, respectively.
        According to \citet{landis77}, the interpretation of a Fliess kappa measure in the range 21--40 per cent is that the raters are in fair agreement.
    %NEXT 2 SENTENCES NEED WORK
        In Fig.~\ref{fig: props_vs_props}, it is clear that the median of properties of classified galaxies using different methods are in agreement. 
        However, since with SOM methods we can have more than 12 groups, the results are more detailed than the other methods.}

>>>>

    In this section we show the results of the neural networks and K-means clustering trained using the \citetalias{Kinney96} template spectra.
    Training with \citetalias{Kinney96} templates results in networks or clusters that have regions corresponding to galaxies of known morphological type. 
    The trained networks or clusters can then be used to categorize other galaxies.

@@

            Figs.~\ref{fig: 1by2T} --~\ref{fig: 1by22T} show the results of the training networks.  
            The upper part of the figures shows the neurons and the relative distances between their weights. 
            As mentioned in Section~\ref{sec: method_somz}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.45\textwidth}
                    \centering
                  \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.45\textwidth}
                    \centering \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption[Results of training network in $1\times2$~grid]{Results of training network in $1\times2$~grid. The upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citet{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

@@

    We performed K-means clustering on the \citetalias{Kinney96} template spectra using 4 clusters.

    Fig.~\ref{fig: som_k_means_4} shows the clustering of the template spectra into 4 groups using both K-means and SOMs.
    Both methods separated the most starburst-like and quiescent galaxies in a similar manner.
    However, they show discrepancies for spectral types between these extreme cases.
    Table~\ref{tab:kmeans_som4} shows the group assignments of the \citetalias{Kinney96} templates in the two methods.
    Templates B, E, S0 and Sa are grouped in the same cluster/neuron in both K-means and $1\times4$-sized SOM. 
    K-means grouped the Sb template with the above templates, but the SOM network classified it as a separate group.
    The K-means method clustered the Sc template as a separate group, while the SOM network grouped it with SB5 and SB6. 
    Another noticeable difference is that K-means clustered the SB1 and SB2 templates in one group and SB3 to SB6 in another while
    the SOM network grouped the SB1 to SB4 templates in one neuron and SB5 and SB6 in another.
    As Fig.~\ref{fig: som_k_means_4} shows, K-means clustering of the templates produced groups more similar to each other, while SOM emphasizes the templates' differences.

@@

@@

         After training with the templates, we used both K-means and self-organizing maps to classify the 142 galaxies in the \citetalias{Hossein12} sample.
         K-means classification was performed by assigned each galaxy to the K-means cluster with the closest centroid and the results are discussed in Section~\ref{sec: Kmeansvssomvsann}.
        This section examines the results of one-dimensional SOM-based classification in detail.

@@

            Since in this network galaxies were forced to be divided into a maximum of two groups, the strongest feature in a galaxy's spectrum predominantly decided to which group the galaxy belonged.
            Galaxies with a weak 4000~\AA~break, strong emission lines and ultraviolet upturn are categorized as starbursts while galaxies with a strong 4000~\AA~break are categorized as quiescent.
            Increasing the size of the SOM helps solve the problem of galaxies which have features that are common in both groups.

            Fig.~\ref{fig: 1by3V} presents the result of classifying the spectra of the galaxies using the $1\times3$~network (from Fig.~\ref{fig: 1by2T}): 66 of the galaxies belong to the quiescent group, and 47 belong to the starburst group. 
            However, 29 of the galaxies are similar to starbursts in some, but not all, of their features. 
            Galaxies in this group have strong emission lines and are ultraviolet-bright, but they also have a strong 4000~\AA~break, which makes them cluster closer to the quiescent galaxies (middle panel in the lower part of Fig.~\ref{fig: 1by2V}).

@@

@@

        As described in Section~\ref{sec: method_somz}, one of the main advantages of the SOM is that when the weight of one neuron is adjusted after finding the best matching unit, the weight of the whole map will be changed.
        This quality of the SOMs provides a unique opportunity to analyse 2D networks using two approaches. 
        At first, as in the 1D SOMs, we assumed that all the galaxies must have spectra similar to any of the 12 templates in Fig.~\ref{fig: k96}~(i.e. no outliers).
        In this case, we can categorize spectra of other galaxies based on those 12 types.
        In the second approach, we trained networks with the assumption that there might be completely different types of galaxies not encountered before.
        This approach can be used to trace and recognize the outliers, which could include galaxy types that are not present in the \citetalias{Kinney96} templates (e.g. dwarf galaxies, luminous infrared galaxies, or active galactic nuclei).
        To execute the second approach, we used the \textsc{selforgmap} library in \textsc{matlab}.
        Unlike the \textsc{newsom} library in which the ordering step neighbourhood distance is set by default (although in an SOM it should be a free parameter), in \textsc{selforgmap} we can change this value.
        Using the \textsc{selforgmap} code, based on the size of a map and the ordering step neighbourhood distance, we can arrange that the data spread all over the SOM or sit close together, allowing room for outliers.

@@

        In the upper left part of the map in Fig.~\ref{fig: 12by12}, although galaxies have more ways to be separated, they were separated into two main groups.
        There is a distinguishable strip of grey, dark grey and in some cases black colour in the map:
        this strip separates quiescent galaxies from starburst ones.
        The lower map shows that 5 of the neurons on the left side of the strip are full. 
        These five neurons are the same as the ones on the left hand side of Fig.~\ref{fig: 1by2T} (quiescent galaxies),
        the only difference being that in this map they have more space to be separated from each other.
        When we use this network to categorize the spectra of galaxies, any galaxy placed on the left hand side of the strip is a quiescent galaxy and any one placed on the right side of the strip is a starburst one.
        The decision about what sub-type of quiescent or starburst galaxy is based upon its position relative to each type in the SOM.
        \citetalias{Hossein12} showed that by masking emission lines (i.e., by ignoring details in spectra) two types of galaxies, Sb and SB6, can be misclassified as one another.
        As mentioned before, few-node, one-dimensional SOMs are a good method to obtain a big picture of the problem under study with less emphasis on fine detail. 
        For example, the SB6 category is placed near Sb galaxies in the 1D network in Fig.~\ref{fig: 1by22T}.
        However, in the upper left panel of Fig.~\ref{fig: 12by12}, we can see that the Sb and SB galaxies are completely separated. 
        This shows that more details of spectral shapes are considered by 2D networks, which is one of the main advantages of  using 2D maps over 1D ones.
       In this case, the two dimensions of the self-organizing map correspond approximately to quiescent versus star-forming (left to right) and less to more extinction (bottom to top).

@@

        We used both 2D networks to categorize the \citetalias{Hossein12} galaxies and the results are shown in the lower panels of Fig.~\ref{fig: 12by12}.
        In the lower right map in Fig.~\ref{fig: 12by12}, all galaxies are placed in the bottom part of the map and we can conclude that in the \citetalias{Hossein12} sample there are no outliers or spectra with very different from the \citetalias{Kinney96} templates.
        The lower left map of Fig.~\ref{fig: 12by12} shows the \citetalias{Hossein12} galaxies categorized based on the network in the upper left of Fig.~\ref{fig: 12by12}. 
        Comparing this categorization with the 1D one from Fig.~\ref{fig: 1by22V}, we can see that only 23 galaxies correspond exactly to \citetalias{Kinney96} types.
        Using 2D maps results in categorizing galaxies into more intermediate types than in 1D maps.
        In both lower maps in Fig.~\ref{fig: 12by12}, most galaxies are in the quiescent side of the SOM, which was predictable from the results in Section~\ref{sec: 1Dv}.
        This does not necessarily imply that in general there are more quiescent galaxies at higher redshift but could be a selection effect for galaxies that had more reliable redshift estimates.

    Although for ease of presentation this paper first discusses 1D networks and continues to 2D networks, we suggest that users of SOMs for galaxy spectral classification should use 2D networks first.
         These can be used as a first step with networks similar to the upper left map in Fig.~\ref{fig: 12by12} to identify outliers in the sample.
        After finding and removing outliers in the input data, the use of 1D and 2D maps is complementary.
        One-dimensional maps are helpful for exploring the general and ordered behaviour of the data, while 2D maps can show a more detailed picture.

\subsection{Comparing classification methods}
    \label{sec: Kmeansvssomvsann}

    The one and two-dimensional self-organizing map classifications are compared briefly above. 
    The 2-D maps allow galaxies to be members of more intermediate classes, and this is a general advantage of SOMs compared to the artificial neural network method used by \citetalias{Hossein12}: in the SOM method galaxies do not need to exactly match template spectra. 
    \citetalias{Hossein12} found that the ANN method could only match 105 out of the 142 galaxies directly to the \citetalias{Kinney96} templates; classifying the remaining 37 galaxies required combining template spectra.
    Below we compare various other classification methods with the results of the 1-D self-organizing maps using the metrics described in Section~\ref{sec:metrics}.
    Because the 2-D SOMs subdivide the sample into a larger number of groups, each containing a small number of objects, they are not directly comparable to the other methods and are not included here.

       To compare results from SOM, K-means, and unsupervised ANN, we first consider how well these methods matched a simple chi-squared fitting classification. 
        Matching galaxy spectra to templates using chi-squared, we found that not all of the templates had matches among the galaxy sample: no galaxy spectra were best-matched to the B, E, or Sc templates. 
        This is consistent with the results from the self-organized maps.
        Summing the classification score over all galaxies in the sample allows us to compare how well each technique matched the results of chi-squared fitting.
        The totals were K-means 124, $1\times4$-sized SOM 103, $1\times12$-sized SOM 111.5, $1\times22$-sized SOM 79, and unsupervised ANN 52.
        This shows that the K-means method and $1\times12$-sized SOM have the most similarity with minimum chi-square fitting.
        Summing the classification score over all methods for a single galaxy shows how well the methods agree for a particular spectrum.
        For example, if the total score of a galaxy is 5, it means that all techniques classified the galaxy into the same group.
        Comparing the scores, we found that 12 per cent of galaxies scored 5, 36 per cent of galaxies had a score greater than 4, and 56 per cent had a total score of 3 or more.
        The score-5 galaxies classified similarly by all methods were primarily (60 per cent) in the Sa template group. 
        The score $>=3$ galaxies were predominantly SB1, Sa, and Sb type (32, 28, 18 per cent, respectively).
        The different methods show the most agreement for galaxies similar to the SB1 and Sa templates, with less agreement on the high-extinction starburst types (SB5, SB6).

       We also used the Fleiss kappa index \citet{landis77} to 
        measure the agreement between methods. 
        Here the `raters' are the different methods classifying a specific number of objects (i.e. 142 galaxies) to a number of classes (i.e. 12 templates).
        We computed two different values of Fleiss kappa, between  K-means, unsupervised ANN, and either 
         $1\times12$-sized or $1\times22$-sized SOMs.
        The resulting values were $\sim 27$ per cent and $\sim 21$ per cent, respectively.
        According to \citet{landis77}, the interpretation of a Fleiss kappa measure in the range 21--40 per cent is that the raters are in fair agreement, consistent with the results above.

A final comparison between the classification methods is 
made with the silhouette score  \citep{rousseeuw87}.
The highest score among our classification was for K-means (0.549), followed by $1\times4$-sized SOM (0.371), $1\times12$-sized SOM (0.302) and $1\times22$-sized SOM (0.249).
The chi-squared fitting resulted in a score of 0.221 and the supervised neural network 0.142.
As discussed in Section~\ref{sec:metrics}, the silhouette score usually declines as the number of groups increases, so we cannot conclude from these values that K-means is the superior method. 
However, the $1\times12$-sized SOM, chi-squared fitting, and supervised neural network all have either 8 or 9 classes.
In this case we can compare scores directly and conclude that the self-organized map yields groups with more internal similarity.
This is in contrast to the comparison between K-means and self-organized maps for the templates alone (see Section\ref{sec: Kmeansvssom}) where K-means clustering  produced groups more similar to each other, and emphasizes that the exact data used in a clustering process can strongly affect the outcome.

 \subsection{Properties of spectral groups}

       To check whether the classifications described above are meaningful, in this section we examine the relations between median properties of the galaxies in each group. 
          Fig.~\ref{fig: props} shows the median of stellar mass, age, specific star formation rate (sSFR; star formation rate per stellar mass), and $f_\mathrm{burst}$ of the galaxies in each neuron in the $1\times2$, $1\times4$, and $1\times8$ self-organized map networks.
        In all plots, the horizontal axis is the number of the neurons divided by the size of the network.
        As shown in Fig.~\ref{fig: props}, in all three networks, stellar mass and age decrease while sSFR and $f_\mathrm{burst}$ increase as the type changes from quiescent galaxies to starburst.
       Separating galaxies based on spectral types also leads to a separation in properties derived (via {\em CIGALE}) from the SEDs, as expected since the spectral types are also based on {\em CIGALE} fitting.

        \begin{figure*}
\includegraphics[width=\textwidth]{images0.01/1d/fig11_new_bw.png}
        \caption{
        From top to bottom: relations between age (t$_{\rm {D4000}}$) and sSFR ($\phi$), sSFR and stellar mass, age (t$_{\rm {D4000}}$) and stellar mass, and A$_{\rm {FUV}}$ with sSFR, age (t$_{\rm {D4000}}$) and stellar mass. 
        Circles, pentagons, and squares show the median values of these properties of galaxies in the $1\times22$, $1\times 12$ and $1\times4$ SOM networks; 
        triangles show the median value of properties for each K-means cluster (using $K=4$) and stars show median values for groups from \citetalias{Hossein12}.
        Error bars show standard deviations of the group properties.
        Colours show the galaxy types in the SOM classification. 
        The lightest grey (most pure red in the online version) colour shows E type galaxies and the black (most blue in the online version) indicates the SB1 galaxies. 
        The shading of other colours between light grey and black (red and blue in the online version) measures the location of galaxies along the SOM from  quiescent to starburst.}
        \label{fig: props_vs_props}
        \end{figure*}

        To compare our classifications with previous work, Fig.~\ref{fig: props_vs_props} shows plots similar to those shown by both \citetalias{Noll09} and \citetalias{Hossein12}.
        The points represent the median values of the properties of the galaxies in each group, with different symbol types representing classification method and error bars indicating the standard deviation of those properties.
       The colour scale represents neuron number in the SOM method, shading from quiescent to starburst. 
        The medium grey (purple in the online version) point in the middle of the black (blue in the online version) points in Fig.~\ref{fig: props_vs_props} corresponds to galaxies with spectra similar to type Sc: the shape of their spectra indicates that they are old galaxies with high sSFR.

        Overall, Fig.~\ref{fig: props} and the upper panels in Fig.~\ref{fig: props_vs_props} show the same trends as noted by \citetalias{Noll09} and \citetalias{Hossein12}:
        galaxies with lower mass tend to be younger and more active.
        The median group properties as derived by all of the methods show high Pearson correlation coefficients ($r^2\gtrsim 0.9$) between the quantities shown.
        The ordering of groups in the SOM maps also corresponds to an ordering in age:    
        from older to younger galaxies, the colour of the points changes from light grey to black (red to blue in the online version).

        \citetalias{Noll09} studied relations between A$_{\rm {FUV}}$ and other properties of galaxies and found that the attenuation had no dependence on the specific star formation or age.
        In contrast to \citetalias{Noll09}, we find a correlation between A$_{\rm {FUV}}$ and these two parameters, shown in the lower panels of Fig.~\ref{fig: props_vs_props}.
        Although weaker then the correlations above, Pearson correlation coefficients are higher for the SOM groups than for other methods, particularly for the $1x12$-sized SOM. 
        The slightly different distribution of this particular classification (see Fig.~\ref{fig: 1by12T}) classifies the most quiescent objects as similar to slightly more active galaxies, lessening the turnover in properties as A$_{\rm {FUV}}$ increases.
